---
title: "Juvenile Snowshoe Hares in Bonanza Creek: Exploratory Report"
author: "Hanna Weyland"
date: "11/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(lubridate)
library(janitor)
library(ggbeeswarm)
library(kableExtra)
library(effsize)
library(broom)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
bonanza_hares <- read.csv(here("data", "bonanza_hares.csv"))
```

### Introduction 

A 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report. You should probably do this last.

```{r fig.cap="AWESOME CAPTION"}

```


### Data and Analysis 

A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used.

Snowshoe hare data was collected.... As keystone species in the northern boreal forests,  

"you will explore the counts and sizes of juvenile snowshoe hares recorded from 1998 - 2012."

### Exploratory Findings 

In this section, complete and communicate the following